Aneuploidy can be an evolutionary detour \\ on the path to adaptation

### 2022-10-30 ms/ms.tex
YP abstract comments

% Abstract 250 words
\begin{abstract} 
Aneuploidy is common in eukaryotes, often leading to decreased growth and fitness. However, evidence from yeast and fungi, as well as human tumour cells, suggests that specific aneuploidies can be beneficial under stressful conditions and facilitate  adaptation.
In a prominent example, an evolutionary experiment with yeast, populations evolving under heat stress had become aneuploid (chromosome III), only to later revert back to euploid after genetic mutations have accumulated. It has therefore been suggested that aneuploidy serves as a "stepping stone" on the path to adaptation.

Here, we test this hypothesis. First, we apply DNA sequencing to show that mutant alleles common in aneuploid cells are uncommon in the evolved euploid population. 
Second, we develop an evolutionary model with both aneuploidy and  mutation, and fit it to the results of the experiment using a Bayesian inference framework.
We then predict the genotype frequency dynamics during the experiment, demonstrating that the majority of the evolved euploid population likely did not descend from aneuploid cells, but rather directly from the euploid wild-type population.
Our model further predicts that if the experiment was repeated with smaller populations, then a larger fraction of the evolved population would descend from aneuploid cells.

Together, our results suggest that aneuploidy can be an inevitable evolutionary "detour" rather than a "stepping stone": it can delay, rather than facilitate, the adaptation of the population, and cells that become aneuploid may leave less descendants compared to cells that remain diploid.
\end{abstract}

>>>>

% Abstract 250 words
\begin{abstract} 
Aneuploidy is common in eukaryotes, often leading to decreased growth and fitness. However, evidence from yeast and other fungi, as well as human tumour cells, suggests that specific aneuploidies can be beneficial under stressful conditions and facilitate  adaptation.
In a prominent example, an evolutionary experiment with yeast, populations evolving under heat stress had become aneuploid (chromosome III), only to later revert back to euploid after genetic mutations have accumulated. It has therefore been suggested that aneuploidy serves as a "stepping stone" on the path to adaptation.

Here, we test this hypothesis. First, we apply DNA sequencing to show that mutant alleles common in aneuploid cells are uncommon in the evolved euploid population. 
Second, we develop an evolutionary model with both aneuploidy and  mutation, and fit it to the results of the experiment using a Bayesian inference framework.
We then predict the genotype frequency dynamics during the experiment, demonstrating that the majority of the evolved euploid population likely did not descend from aneuploid cells, but rather directly from the euploid wild-type population.
%Our model further predicts that if the experiment was repeated with smaller populations, then a larger fraction of the evolved population would descend from aneuploid cells.
Our model shows how the beneficial mutation supply--the product of population size and beneficial mutation rate--determine the evolutionary dynamics, such that with a lower mutation supply a larger fraction of the evolved population may descend from aneuploid cells. However, with high mutation supply, beneficial mutations are generated before fixation of aneuploidy, and can outcompete aneuploidy due to its inherent fitness cost.

Together, our results suggest that aneuploidy can be an evolutionary "detour" rather than a "stepping stone": it can delay, rather than facilitate, the adaptation of the population, and cells that become aneuploid may leave less descendants compared to cells that remain diploid.
\end{abstract}
